---
title: "34.1 - Linear regression notes"
output: github_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
# figs folder
if (!file.exists("figs")) {
  dir.create("figs")
}
# chunk options
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/", # location of files
  fig.height = 5.5, # height of figures
  fig.width = 8 # width of figures
) # width of figures
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = "1000",
  scipen = 100000000
)
options(max.print = 999999)
```

Load packages below. 

```{r packages, message=FALSE, warning=FALSE}
library(magrittr) # Pipes %>%, %T>% and equals(), extract().
library(tidyverse) # all tidyverse packages
library(fs) # file management functions
library(fivethirtyeight)
library(mdsr) # modern data science with R
library(broom) # tidying models
library(modelr) # modelr package
library(ggrepel) # repel text overlay in graphs
library(gridExtra) # arrange multiple plots
library(grid) # text for graph title
library(egg) # for ggarrange
library(corrr) # correlations
library(skimr) # summaries 
library(hrbrthemes) # themes for graphs
library(moderndive) 
library(gapminder)
```

## From 

> ...try to explain diﬀerences in instructor teaching scores as a function of one numerical variable:

> the instructor's “beauty score”; we'll describe how this score was computed shortly. Could it be that instructors with higher beauty scores also have higher teaching evaluations? Could it be instead that instructors with higher beauty scores tend to have lower teaching evaluations? Or could it be there is no relationship between beauty score and teaching evaluations? We'll answer these questions by modeling the relationship between teaching scores and “beauty scores” using simple linear regression where we have:

> 1. A numerical outcome variable `y`, the instructor's teaching score and

> 2. A single numerical explanatory variable `x`, the instructor's beauty

```{r EvalsChp6}
EvalsChp6 <- moderndive::evals %>% dplyr::select(ID, 
                                     score, 
                                     bty_avg, 
                                     age)
EvalsChp6 %>% glimpse(78)
```

## EDA 

> 1. Most crucially: Looking at the raw data values.  
> 2. Computing summary statistics, like means, medians, and interquartile ranges.  
> 3. Creating data visualizations.  
> In it is important to note that the observational unit are individual courses and not individual instructors.

```{r score_model}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = EvalsChp6) # Get regression table: 
get_regression_table(score_model)
```

> For every increase of `1` unit in `bty_avg`, there is an associated increase of, on average, `0.0666` units of `score`.



```{r regression_points}
regression_points <- get_regression_points(score_model) 
regression_points %>% head()
```

