---
title: "Storybench Drafts"
output: 
  github_document: 
    df_print: kable
    toc: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
library(hrbrthemes)
# figs folder
if (!file.exists("figs")) {
  dir.create("figs")
}
# chunk options
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/", # location of files
  fig.height = 7.5, # height of figures
  fig.width = 10 # width of figures
) # width of figures
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
```

This repo is for working drafts on Storybench's [Data Journalism in R](https://www.storybench.org/category/data-journalism-in-r/). 

The .Rmd files are listed below: 

```{r Rmd-files}
fs::dir_tree(".", regex = ".Rmd")
```

When these are complete, they get added to the [storybench-posts repo](https://github.com/mjfrigaard/storybench-posts). 

## 04.1-manipulate-variables-with-dplyr

## 04.2-manipulate-cases-with-dplyr

## 05.1-ggplot2-qplot

## 05.2-ggplot2-extras

## 25.1-tidydata-group-pivot-long-wide

## 34.0-linear-regression-assumptions

## 34.1-linear-regression-models

## 34.9-linear-regression-notes




